---
title: "Project 3"
author: "Qinzhi Peng"
date: "Fall 2022"
output: html_document
---

```{r message=FALSE, warning=FALSE, echo=FALSE}
# load libraries
library(tidyverse)
library(corrplot)
library(GGally)
library(bestglm)
library(pROC)
library(FNN)
library(e1071)
library(rpart)
library(randomForest)

# read csv file and convert it to data frame
df <-  data.frame(read.csv("wineQuality.csv", head = TRUE, stringsAsFactors = TRUE))
```

## Exploratory Data Anaylsis

### Basic Information
#### Observations

First, We can see a few observations below. It seems that there is no obviously non-informative column.

```{r}
# see a few observations
head(df)
```

#### Data Description

Get some basic information for this new data set. 

```{r}
str(df)
```

There are 6497 rows and 12 columns in the data set. It contains 11 predictor variables - all are quantitative, and a response variable that is `label`.

### Data Distributions
#### Response variable

Show distributions of the response variable `label`.

```{r}
ggplot(data = df, mapping=aes(x = label)) +
  geom_bar(fill="blue") + 
  ggtitle("Label Distribution")
```

We can see that two classes are unbalanced, so we should not assume a class-separation threshold of 0.5.

#### Predictor Variables

```{r}
df.qnt <- df %>% 
  dplyr::select(-label) %>%
  gather()

#  visualize the data in multiple columns, side by side
ggplot(data = df.qnt, mapping = aes(x = value)) +
  geom_histogram(color = "blue", fill = "yellow", bins = 30) +
  facet_wrap(~ key, scales = 'free') 
```

We can see there are a few variables are highly right-skewed, like `free.sd`, `sugar`. So we will do the transformation for these variables.

####  response vs. predictor variables

Show grouped box plots of `label` versus all the predictor variables for the log data set.

```{r}
num.qnt <- 11

ggplot(data = df.qnt, mapping = aes(x = rep(df$label, num.qnt), y = value)) +
  geom_boxplot(color="red", fill="blue") +
  facet_wrap(~ key, scales = 'free_y') +
  ylab("price")
```

We can see there are some clear associations between the `label` and predictor variables. For example, with a great percentage of alcohol, the `label` is more likely to be GOOD. That implies we can produce some statistical model to predict `label` with these variables. 


### Data Processing
#### Summary

Get a textual summary of the data set and it seems that there are no missed values or weird values in the data set.

```{r}
summary(df)
```

#### Data Transformation

Do the log transformation for highly skewed variables.

```{r}
# define a new data set for the log transformation
df.log <- df
df.log$vol.acid = log(df.log$vol.acid)
df.log$chlorides = log(df.log$chlorides)
df.log$free.sd = log(df.log$free.sd)
df.log$sulphates = log(df.log$sulphates)
df.log$sugar = log(df.log$sugar)

df.qnt.log <- df.log %>% 
  dplyr::select(-label) %>%
  gather()

#  visualize the data in multiple columns, side by side
ggplot(data = df.qnt.log, mapping = aes(x = value)) +
  geom_histogram(color = "blue", fill = "yellow", bins = 30) +
  facet_wrap(~ key, scales = 'free') 
```

#### Removing Outliers

As we can see in the figures, there are some obvious outliers in the data set. 

```{r}
df.new <- df.log
# examine outliers
df[which(df.log$citric > 1),]
df[which(df.log$total.sd > 350),]

# define the indexes of outliers
index.outiers <-  which(df.log$citric > 1 | df.log$total.sd > 350)

# define new data set
df.new <- df.log[-index.outiers,]
```

These outliers have really large `citric` and `total.sd` values, we think these observations maybe are some special cases. Then, we remove all these outliers before training our model.

```{r}
# visualize the data after the data processing
df.qnt.new <- df.new %>% 
  dplyr::select(-label) %>%
  gather()

ggplot(data = df.qnt.new, mapping = aes(x = value)) +
  geom_histogram(color = "blue", fill = "yellow", bins = 30) +
  facet_wrap(~ key, scales = 'free') 
```
After the log transformation and the removal of outliers, we can see that the distributions predictors appear to not be highly skewed.

### Correlation

```{r}
df.new %>% 
  dplyr::select(-label) %>%
  ggpairs(lower = list(combo = wrap("facethist", binwidth = 0.8)), progress=FALSE)
```

We can clearly see the correlation coefficients between the `total.sd` and `free.sd` is all greater than 0.7, which means they are correlated. Since our goal is prediction, the multicollinearity is acceptable in this case.


## Model Selection
### Split the data 

Split the data into training and test sets in a 70%-30% ratio. 

```{r}
# split the response and predictors
response   <- df.new[,12]  
predictors <- data.frame(scale(df.new[,-12]))

# split training and test
set.seed(100)
s <- sample(nrow(df.new), round(0.7*nrow(df.new)))
pred.train <- predictors[s,]
pred.test <- predictors[-s,]
resp.train <- response[s]
resp.test <- response[-s]
df.train <- cbind(pred.train,resp.train)
df.test <- cbind(pred.test,resp.test)
```

### Logistic Regression

First, we do the Logistic Regression between response and predictor variables. Also, we output Class 1 probabilities and record its ROC result.

```{r}
out.glm = glm(resp.train ~ ., data = pred.train, family = binomial)
# extract Class 1 probabilities
resp.prob.glm = predict(out.glm, newdata = pred.test, type = "response")  
# output a few of probabilities
resp.prob.glm[1:10]
```

```{r message=FALSE}
roc.glm = roc(resp.test, resp.prob.glm)
```

Meanwhile, we perform the Best-subset-selection for Logistic model, and use the best model to get Class 1 probabilities.

```{r message=FALSE}
# expect the response variable to be named y
y <- df.train$resp.train
df.train.bestglm <- df.train[,-12]
df.train.bestglm <- data.frame(df.train.bestglm,"y"=y)
y <- df.test$resp.test
df.test.bestglm <- df.test[,-12]
df.test.bestglm <- data.frame(df.test.bestglm,"y"=y)

# variable selection
out.aic <- bestglm(df.train.bestglm, family = binomial, IC = "AIC")
out.aic$BestModel

# get class 1 prob
resp.prob.bestglm = predict(out.aic$BestModel, newdata = df.test.bestglm, type = "response") 
```

For the AIC, there are 7 predictor variables retained in the best models, which are `vol.acid`, `sugar`, `free.sd`, `total.sd`, `density`, `sulphates`, and `alcohol`.

```{r message=FALSE}
resp.prob.bestglm[1:10]

roc.bestglm = roc(resp.test, resp.prob.bestglm)
```

### Classification Trees
```{r}
out.rpart = rpart(resp.train ~ .,data = pred.train)
# extract Class 1 probabilities
resp.prob.rpart = predict(out.rpart, newdata = pred.test, type="prob")[,2]
resp.prob.rpart[1:10]
```

```{r message=FALSE}
roc.rpart = roc(resp.test, resp.prob.rpart)
```

### Random Forest

```{r message=FALSE}
out.rf = randomForest(resp.train ~ ., data = pred.train, importance=TRUE)
# extract Class 1 probabilities
resp.prob.rf = predict(out.rf, newdata = pred.test, type="prob")[,2]
resp.prob.rf[1:10]
```

```{r message=FALSE}
roc.rf = roc(resp.test, resp.prob.rf)
```

### KNN

For the KNN model, we need to first choose the best `k` value.

```{r}
k.max = 30
mcr.k = rep(NA,k.max)
for ( kk in 1:k.max ) {
  knn.cv = knn.cv(train = pred.train, cl = resp.train, k = kk) 
  mcr.k[kk] = mean(knn.cv!=resp.train)
}
k.min = which.min(mcr.k)
cat("The optimal number of nearest neighbors is ",k.min,"\n")
```

Plot the validation-set MCR versus k.

```{r}
# plot the validation-set MCR versus k
ggplot(data=data.frame("k"=1:k.max, "mcr"=mcr.k), mapping=aes(x=k, y=mcr)) + 
  geom_point() + geom_line() +
  xlab("Number of Nearest Neighbors k") + ylab("Validation MCR") + 
  geom_vline(xintercept=k.min,color="red")
```

```{r}
out.knn = knn(train = pred.train, test = pred.test, cl = resp.train, k = k.min, algorithm="brute", prob=TRUE)
# get class 1 probabilities
resp.prob.knn = attributes(out.knn)$prob
w = which(out.knn=="BAD") 
resp.prob.knn[w] = 1 - resp.prob.knn[w]
# output a few probs
resp.prob.knn[1:10]
```

```{r message=FALSE}
roc.knn = roc(resp.test, resp.prob.knn)
```

### SVM

For the SVM model, it will take a long time to determine an optimal `cost` value. Thus, We just assume `cost = 1` here, which might not be optimal but save time. We also try both linear and polynomial kernels in the SVM models.

```{r warning=FALSE}
# linear kernel
out.svm.linear  = tune(svm, resp.train ~ . , data = df.train, kernel="linear", ranges=list(cost=1), prob = TRUE)
resp.pred.svm.linear = predict(out.svm.linear$best.model, newdata=df.test, prob = TRUE)
resp.prob.svm.linear = attributes(resp.pred.svm.linear)$prob[, 1]
# show a few probs
resp.prob.svm.linear[1:10]

# polynomial kernel
out.svm.poly = tune(svm, resp.train ~ ., data = df.train, kernel="polynomial", ranges=list(cost=1, degree=2:4), prob = TRUE)
resp.pred.svm.poly = predict(out.svm.poly$best.model, newdata=df.test, prob = TRUE)
resp.prob.svm.poly = attributes(resp.pred.svm.poly)$prob[, 1]
# show a few probs
resp.prob.svm.poly[1:10]
```

```{r message=FALSE}
roc.svm.linear = roc(resp.test, resp.prob.svm.linear)
roc.svm.ploy = roc(resp.test, resp.prob.svm.poly)
```

### ROC curves & AUC

Plot and compare ROC curves for different model, and output their AUC values.

```{r message=FALSE}
ggroc(list(LG = roc.glm, BestGLM = roc.bestglm, CT = roc.rpart, RF = roc.rf, KNN = roc.knn, SVM.linear = roc.svm.linear, SVM.ploy = roc.svm.ploy),) + theme_light() + ggtitle("ROC Curves")
```


```{r}
knitr::kable(data.frame(roc.glm$auc, roc.bestglm$auc, roc.rpart$auc, roc.rf$auc, roc.knn$auc, roc.svm.linear$auc, roc.svm.ploy$auc))
```

We can see clearly that the Random Forest model has the highest AUC value. So we use this model to make class predictions.

## Prediction
### class-separation Threshold

Use Youden’s J statistic to determine the optimal class-separation threshold. We assume that sensitivity and specificity are equally important.

```{r message=FALSE}
# find the threshold associated with that maximum value
optimal.idx <- which.max(roc.rf$sensitivities + roc.rf$specificities - 1)
# the optimal class-separation threshold
threshold <- roc.rf$thresholds[optimal.idx]
threshold
```

The optimal threshold is 0.611, which approaches the proportion of two classes. Then, using the threshold, we make class predictions for the Random Forest model.

### Confusion Matrix

```{r}
# class predictions
resp.pred.rf <- as.factor(ifelse(resp.prob.rf > threshold, "GOOD", "BAD"))
# create a confusion matrix 
tab.rf <- table(resp.pred.rf, resp.test)
tab.rf
```

From the confusion matrix, we can see we will get more false negatives than false positives in the prediction.

### Misclassification rate

```{r}
misrate.test <- 1-sum(diag(tab.rf))/sum(tab.rf)
misrate.test
```

The misclassification rate is 0.18. That means our model is good for predictions and also has relatively high sensitivity and specificity. 
